Finch: Sparse and Structured Array Programming with Control Flow

### 2024-04-05 sections/eval/graphs.tex
need to edit contributions
Update on Overleaf.

We used Finch to implement two fundamental graph algorithms: Breadth-first search (BFS) and Bellman-Ford single-source shortest path. Our BFS implementation and graphs datasets are taken from Yang et al.~\cite{yang_implementing_2018}, including road networks, characterized by bounded node degrees and long diameters, and scale-free graphs, where node degree distribution follows a power-law distribution and diameters are short.

Direction-optimization~\cite{beamer2012direction} is crucial for achieving high BFS performance in such scenarios, switching between push and pull traversals to efficiently explore graphs. Push traversal visits the neighbors of each frontier node, while pull traversal visits every node and checks to see if it has a neighbor in the frontier. The advantage of pull traversal is that we may terminate our search once we find a node in the frontier, saving time in the event the push traversal were to visit most of the graph anyway. Early break is the critical part of control flow in this algorithm, though the algorithms also require different loop orders, multiple outputs, and custom operators.

Figure~\ref{fig:graph_result} compares performance to Graphs.jl, an open-source Julia library, and the LAGraph Library, which implements graph algorithms in the language of linear algebra using GraphBLAS~\cite{mattson2019lagraph}. For the BFS algorithm, direction-optimization notably enhances performance for scale-free graphs. Although GraphBLAS uses hardwired optimizations, Finch is competitive on these benchmarks. On Bellman-Ford (with path lengths and shortest-path tree), Finch's support for multiple outputs, sparse inputs, and masks leads to superior performance over GraphBLAS (average speedup of 3.53). Note that we did not include GAP-road as it took too long to run. %speedup is arith-mean

>>>>

We used Finch to implement both Breadth-first search (BFS) and Bellman-Ford single-source shortest path.
%
Our BFS implementation and graphs datasets are taken from Yang et al.~\cite{yang_implementing_2018}, including both road networks and scale-free graphs (bounded node degree vs. power law node degree).

%road networks, characterized by bounded node degrees and long diameters, and scale-free graphs, where node degree distribution follows a power-law distribution and diameters are short.

Direction-optimization~\cite{beamer2012direction} is crucial for achieving high BFS performance in such scenarios, switching between push and pull traversals to efficiently explore graphs.
%
Push traversal visits the neighbors of each frontier node, while pull traversal visits every node and checks to see if it has a neighbor in the frontier. 
%
The advantage of pull traversal is that we may terminate our search once we find a node in the frontier, saving time in the event the push traversal were to visit most of the graph anyway. 
%
Early break is the critical part of control flow in this algorithm, though the algorithms also require different loop orders, multiple outputs, and custom operators.

Figure~\ref{fig:graph_result} compares performance to Graphs.jl, a Julia library, and the LAGraph Library, which implements graph algorithms with sparse linear algebra using GraphBLAS~\cite{mattson2019lagraph}.
%
For the BFS algorithm, direction-optimization notably enhances performance for scale-free graphs. 
%
Although GraphBLAS uses hardwired optimizations, Finch is competitive. 
%
On Bellman-Ford (with path lengths and shortest-path tree), Finch's support for multiple outputs, sparse inputs, and masks leads to superior performance over GraphBLAS (average speedup of 3.53). 
%
We did not include GAP-road as it timed out. %speedup is arith-mean
%
In Appendix~\ref{sec:graph_listings}, we display the code for BFS and Bellman-Ford in Finch (57 and 50 LOC) and LAGraph (215 and 227 LOC), and invite readers to compare the clarity of the algorithms.

### 2024-04-05 sections/eval/morphology.tex
Update on Overleaf.

Some image processing pipelines stand to benefit from structured data processing \cite{donenfeld_unified_2022}.
In this case study, we focus on binary image morpology and the logical transformation of binary images and masks.
We consider two operations: binary erosion (computing a mask), and a masked histogram (using a mask to avoid work).

Samples from our datasets are shown in Figure~\ref{fig:image_datasets}. Note that these images are all binary, either by design or having been thresholded.

Finch allows us to modify both our algorithm and our datastructure, so we may choose to use either a dense representation with bytes ($Dense(Element(0x00))$), a bit-packed representation ($Dense(Element(UInt64))$), or a run-length encoded representation that represents runs of true or false ($SparseRLE(Pattern())$).
%
All of these have their advantages.
%
The dense representation induces the least overhead, the bit-packed representation can take advantage of bitwise binary ops, and the run-length encoded version only uses memory and compute when the pattern changes between true and false.

The erosion operation turns off a pixel unless all of it's neighbors are on.
%
This can be used to shrink the boundaries of a mask, and remove point instances of noise \cite{fisher_hypermedia_1996}.
\help{link a few cool images that show erosion}
%
Note that three instances of structured data are used here, both in the mask, in the padding of the inputs, and in the convolutional filter itself.
%
We might understand the filter as a structured tensor of circular shifts, or we might understand each shifted view of the data in an unrolled stencil computation as a structured tensor.
Figure \ref{fig:erosion_listing} displays example erosion algorithms for bitwise
or run-length-encoded algorithms, and a masked histogram kernel.

We compare to OpenCV, and chose the histogram since the OpenCV histogram function also accepts a mask. If we use $SparseRLE(Pattern())$ for our mask, we can reduce the branching
in the masked kernel and get better performance.
%
We this is also an instance where we use a computed index into the output, something not many sparse and structured programming frameworks support.

We used four datasets. We randomly selected 100 images from the MNIST \cite{lecun_gradient-based_1998} and Omniglot \cite{lake_human-level_2015} character recognition datasets, as well as a dataset of human line drawings \cite{eitz_how_2012}. We also hand-selected a subset of mask images (these images were less homogenous, so we listed them in the appendix) from a digital image processing textbook \cite{gonzalez_digital_2006}. All images were thresholded, and we also include versions of the images that have been magnified before thresholding, to induce larger constant regions. IN our erosion task, the sparseRLE format performs the best as it is asymptotically faster and uses less memory, leading to a 19.5X speedup over OpenCV on the sketches dataset, which becomes arbitrarily large as we magnify the images (here shown as 266X). The improvements with SparseRLE are seen in the histogram task as well, as it allows us to
mask off contiguous regions of computation, instead of individual pixels, reducing the branches in the mask and leading to a significant speedup (20.3x on Omniglot and 20.8x on sketches. We believe the 51.6x on MNIST is due to calling overhead in OpenCV). The bitwise kernels were effective as well, and would be more effective on datasets with less structure. A strength of Finch is that it supports structured datasets, even over bitwise operations, allowing us to implement the bitwise kernel and then mask it.

>>>>

Some image processing pipelines stand to benefit from structured data processing \cite{donenfeld_unified_2022}.
%
We focus on binary image morpology and the logical transformation of binary images and masks.
%
We consider two operations: binary erosion (computing a mask), and a masked histogram (using a mask to avoid work).
%
We use images are all binary, either by design or having been thresholded.

Finch allows us choose our datastructure, so we may choose to use either a dense representation with bytes ($Dense(Element(0x00))$), a bit-packed representation ($Dense(Element(UInt64))$), or a run-length encoded representation that represents runs of true or false ($SparseRLE(Pattern())$).
%
All of these have their advantages.
%
The dense representation induces the least overhead, the bit-packed representation can take advantage of bitwise binary ops, and the run-length encoded version only uses memory and compute when the pattern changes. % between true and false.

Similarly, since Finch let's us choose our algorithm we can implement erosion in a few ways.
%
The erosion operation turns off a pixel unless all of it's neighbors are on.
%
This can be used to shrink the boundaries of a mask, and remove point instances of noise \cite{fisher_hypermedia_1996}.
%\help{link a few cool images that show erosion}
%
This introduces three instances of structure in the control flow: the mask, the padding of inputs, and the convolutional filter.
%
We focused on the filter.
%Note that three instances of structured data are used here, both in the mask, in the padding of the inputs, and in the convolutional filter itself.
%
We might understand the filter as a structured tensor of circular shifts, or we might understand each shifted view of the data in an unrolled stencil computation as a structured tensor, or a two part stencil where we compute the horizontal then vertical part of the stencil.
%
We experimented with these options and found that the last approach performed best, due to fitting the storage formats while reducing the amount of work with intermediate temporaries.
%
Figure \ref{fig:morphology_listing} displays example erosion algorithms for bitwise
or run-length-encoded algorithms.
%
% and a masked histogram kernel.

We compared against OpenCV.
%
We used four datasets. We randomly selected 100 images from the MNIST \cite{lecun_gradient-based_1998} and Omniglot \cite{lake_human-level_2015} character recognition datasets, as well as a dataset of human line drawings \cite{eitz_how_2012}. 
%
We also hand-selected a subset of mask images (these images were less homogenous, so we listed them in the appendix) from a digital image processing textbook \cite{gonzalez_digital_2006}. 
%
All images were thresholded, and we also include versions of the images that have been magnified before thresholding, to induce larger constant regions. 
%
IN our erosion task, the sparseRLE format performs the best as it is asymptotically faster and uses less memory, leading to a 19.5X speedup over OpenCV on the sketches dataset, which becomes arbitrarily large as we magnify the images (here shown as 266X). 
%
We believe the 51.6x on MNIST is due to calling overhead in OpenCV). 
%
The bitwise kernels were effective as well, and would be more effective on datasets with less structure. 
%
A strength of Finch is that it supports structured datasets, even over bitwise operations, allowing us to implement the bitwise kernel and then mask it.

We also implemented a histogram kernel.
%
We used an indirect access into the output to implement this (Figure \ref{fig:morphology_listing}), something not many sparse frameworks support.
%
We compare to OpenCV since the OpenCV histogram function also accepts a mask. 
%
If we use $SparseRLE(Pattern())$ for our mask, we can reduce the branching
in the masked kernel and get better performance.
%
The improvements with SparseRLE are seen in the histogram task too, as it allows us to mask off contiguous regions of computation, instead of individual pixels, reducing the branches and leading to a significant speedup (20.3x on Omniglot and 20.8x on sketches).
%
%
%A strength of Finch is that it supports structured datasets, even over bitwise operations, allowing us to implement the bitwise kernel and then mask it.

### 2024-04-05 sections/eval/spgemm.tex
Update on Overleaf.

Sparse Matrix-Matrix Multiplication (SpGEMM) is a fundamental operation in scientific computing and data analytics. 
We compute the $M \times N$ sparse matrix $C$ as the product of $M \times K$ and $K \times N$ sparse matrices $A$ and $B$.

There are three main approaches to SpGEMM \cite[Section 2.2]{zhang2021gamma}.
%
The inner-products algorithm takes dot products of corresponding rows and columns, while the outer-products algorithm sums the outer products of corresponding columns and rows.
%
Gustavson's algorithm sums the rows of $B$ scaled by the corresponding nonzero columns in each row of $A$.
%
Inner-products is known to be asymptotically less efficient than the others, as we must do a merge operation to compute each of the $O(MN)$ entries in the output \cite{ahrens2022autoscheduling}.

Figure~\ref{fig:spgemm_listing} implements all three approaches in Finch, and Figure~\ref{fig:spgemm} compares the performance of Finch to TACO.
%
Note that these algorithms mainly differ in their loop order, but that different datastructures can be used to support the various access patterns induced.
%
Although a sparse bytemap has a dense memory footprint, we use it in our Finch implemenation of Gustavson's for the smaller $O(N)$ intermediate.
%
In our Finch implementation of outer products, we use a sparse hash table, as it is fully-sparse and randomly accessible.
%
Unfortunately, TACO does not support multidimensional sparse workspaces, only supporting outer-products with a dense output, a limitation with asymptotic consequences.
%
It is also worth mentioning that the bytemap format in TACO's Gustavson's implementation is hard-wired, whereas Finch's programming model allows us to write algorithms with explicit temporary formats and transpositions.

As depicted in Figure~\ref{fig:spgemm}, Finch achieves comparable performance with TACO on smaller matrices when we use the same datastructures, and significant improvements when we use better datastructures. Finch outperforms TACO on larger matrices, with an average speedup of 1.25. %arith-mean

>>>>

%Sparse Matrix-Matrix Multiplication (SpGEMM) is a fundamental operation in scientific computing and data analytics. 
We compute the $M \times N$ sparse matrix $C$ as the product of $M \times K$ and $K \times N$ sparse matrices $A$ and $B$.
%
There are three main approaches to SpGEMM \cite[Section 2.2]{zhang2021gamma}.
%
The inner-products algorithm takes dot products of corresponding rows and columns, while the outer-products algorithm sums the outer products of corresponding columns and rows.
%
Gustavson's algorithm sums the rows of $B$ scaled by the corresponding nonzero columns in each row of $A$.
%
Inner-products is known to be asymptotically less efficient than the others, as we must do a merge operation to compute each of the $O(MN)$ entries in the output \cite{ahrens2022autoscheduling}.
%
We will show that our ability to implement these latter methods exceeds that of TACO, translating to asymptotic benefits. 
%

Figure~\ref{fig:spgemm_listing} implements all three approaches in Finch, and Figure~\ref{fig:spgemm} compares the performance of Finch to TACO.
%
Note that these algorithms mainly differ in their loop order, but that different data structures can be used to support the various access patterns induced.
%
In our Finch implementation of outer products, we use a sparse hash table, as it is fully-sparse and randomly accessible.
%
Since, TACO does not support multidimensional sparse workspaces, its outer products uses a dense intermediate, which leads to an asymptotic slow down shown in Figure~\ref{fig:spgemm}.
%
Similarly, although a sparse bytemap has a dense memory footprint, we use it in our Finch implementation of Gustavson's for the smaller $O(N)$ intermediate.
%
We note that the bytemap format in TACO's Gustavson's implementation is hard-wired, whereas Finch's programming model allows us to write algorithms with explicit temporaries and transpositions.
%
Without such hard wiring, TACO would have to use a dense intermediate to support random writes.
%
But without if conditions, TACO would then be forced to propagate all values in the dense vector to the sparse output matrix, turning the output dense and leading to the same asymptotic results as in the case of outer products. 
%
As depicted in Figure~\ref{fig:spgemm}, Finch achieves comparable performance with TACO on smaller matrices when we use the same datastructures, and significant improvements when we use better datastructures. 
%
Finch outperforms TACO on larger matrices, with an average speedup of 1.25. %arith-mean

### 2024-04-05 sections/eval/spmv.tex
Update on Overleaf.

\begin{wrapfigure}{r}{0.41\linewidth}
    \vspace{-28pt}
    \begin{minipage}[t]{0.18\textwidth}
        \begin{minted}{julia}
            y .= 0
            for j = _, i = _
              y[i] += A[i, j] * x[j]
            end
        \end{minted}
        \vspace{24pt} % Add this to ensure top alignment within minipage
        \begin{minted}{julia}
            y .= 0
            for j = _, i = _
              y[j] += A[i, j] * x[i]
            end
        \end{minted}
    \end{minipage}\hfill%
    \begin{minipage}[t]{0.22\textwidth}
        \vspace{0pt} % Add this to ensure top alignment within minipage
        \begin{minted}{julia}
            y .= 0
            for j = _
              let x_j = x[j]
                y_j .= 0
                for i = _
                  let A_ij = A[i, j]
                    y[i] += x_j * A_ij
                    y_j[] += A_ij * x[i]
                  end
                end
                #D is the diagonal
                y[j] += y_j[] + D[j] * x_j
              end
            end
        \end{minted}
    \end{minipage}
    \vspace{-8pt}
    \caption{Finch row-major, column-major and symmetric SpMV Programs}
    \label{spmv_programs}
    \vspace{-12pt}
  \end{wrapfigure}
  \subsection{Sparse Matrix-Vector Multiply (SpMV)}
  Sparse matrix-vector multiplication (SpMV) is perhaps the most studied sparse
  kernel, with a wide range of applications \cite{liu_csr5_2015,
  zhou_enabling_2020}. Because SpMV is a bandwidth bound kernel, many formats have
  been proposed to reduce the footprint with a minimal impact on complexity
  \cite{langr_evaluation_2016}. The wide range of applications unsurprisingly
  results in a wide range of array structures, making it an effective kernel to
  demonstrate the utility of our programming model.

  We varied both the data formats and the programs we used them with. Our formats
  are shown in Table~\ref{spmv_tensor_formats}, and our programs are listed in
  Figure~\ref{spmv_programs}.

  In varying the program, we consider both row and column-major SpMV programs, as
  well as a symmetric variant of SpMV. Finch enables us to exploit symmetry
  effectively. Our program restricts our attention to the canonical trial (using
  masks), reuses the reads to the canonical triangle of the symmetric matrix
  (using a $\finchdefine$ statement), and writes the results to both relevant
  locations (using multiple outputs).

  All of our programs apply to all of our level formats, enabling exploitation of multiple structural patterns concurrently (e.g. sparsity \textit{and} symmetry).

  Figure~\ref{fig:spmv_grouped} displays speedup relative to TACO, SuiteSparseGraphBLAS, and Julia’s standard library.  We test using sparse matrices from a large selection of datasets spanning several previous papers: the datasets used by Vuduc et al. to test the OSKI interface \cite{vuduc2005oski}, Ahrens et al. to test a variable block row format partitioning strategy \cite{ahrens_optimal_2021}, and Kjolstad et al. to test the TACO library \cite{kjolstad_tensor_2017}. Additionally, we included the SNAP graph collection to test with boolean matrices. We also created some synthetic matrices containing bands of varying sizes.
  % or blocks of varying sizes as well as a permutation matrix to encapsulate a few additional use cases.
  The dense vector is randomly generated. We tested using the row-major and column-major Finch programs in Figure \ref{spmv_programs} as well as the symmetric program where applicable; the performance displayed for Finch on each dataset is the fastest among the formats and programs we tested. Column-major SpMV consistently performs better than row-major SpMV (an average of 1.36x better) in TACO so we use column-major SpMV in TACO as our baseline.

  We found that the SpMV performance was superior for the level format that best paralleled the structure of the tensor.
  Namely, matrices with a clear blocked structure like exdata\_1, TSOPF\_RS\_b678\_c1, and heart3 performed notably well with the SparseVBL format with speedups of 2.16, 1.55, and 1.30 relative to TACO, while the baseline format had slowdowns of 0.71, 0.53, and 0.92 relative to TACO. Furthermore, the synthetic banded matrices we constructed performed the best with the SparseBand matrix, in particular with the large\_band and the medium\_band matrices having a speedup of 1.98 and 1.64 relative to TACO, while the baseline format had slowdowns of 0.84 and 0.51 relative to TACO.
  There were also significant advantages of using the Pattern format instead of the Element format to represent scalar values in the matrices when these values were boolean, such as matrices in the SNAP collection which represent graph datasets are boolean. For example, the SparseList-Pattern for email-Eu-core resulted in a speedup of 2.51, while the SparseList-Element format resulted in a slowdown of 1.84 over TACO.
  Every symmetric matrix in the SparseList and SparseList-Pattern formats has better performance when we use a Finch SpMV program that takes advantage of this symmetry.
  Symmetric SpMV with the SparseList level format in Finch results in an average of 1.27x speedup over TACO and symmetric SpMV with the SparseList-Pattern format in Finch results in an average speedup of 1.21x over TACO.
  Notably, there is a 1.91x speedup for the HB/saylr4 matrix over TACO.

@@

  \begin{wrapfigure}{r}{0.41\linewidth}
      \begin{minipage}[t]{0.18\textwidth}
          \vspace{0pt} % Add this to ensure top alignment within minipage
          \begin{minted}{julia}
              y .= 0
              for j = _, i = _
                y[i] += A[i, j] * x[j]
              end
          \end{minted}
          \vspace{24pt} % Add this to ensure top alignment within minipage
          \begin{minted}{julia}
              y .= 0
              for j = _, i = _
                y[j] += A[i, j] * x[i]
              end
          \end{minted}
      \end{minipage}\hfill%
      \begin{minipage}[t]{0.22\textwidth}
          \vspace{0pt} % Add this to ensure top alignment within minipage
          \begin{minted}{julia}
              y .= 0
              for j = _
                let x_j = x[j]
                  y_j .= 0
                  for i = _
                    let A_ij = A[i, j]
                      y[i] += x_j * A_ij
                      y_j[] += A_ij * x[i]
                    end
                  end
                  #D is the diagonal
                  y[j] += y_j[] + D[j] * x_j
                end
              end
          \end{minted}
      \end{minipage}
      \vspace{-8pt}
      \caption{Finch row-major, column-major and symmetric SpMV Programs}
      \label{spmv_programs}
      \vspace{-12pt}
  \end{wrapfigure}

>>>>

\begin{wrapfigure}{r}{0.41\linewidth}
    \vspace{-28pt}
    \begin{minipage}[t]{0.18\textwidth}
        \begin{minted}{julia}
            y .= 0
            for j = _, i = _
              y[i] += A[i, j] * x[j]
            end
        \end{minted}
        \vspace{24pt} % Add this to ensure top alignment within minipage
        \begin{minted}{julia}
            y .= 0
            for j = _, i = _
              y[j] += A[i, j] * x[i]
            end
        \end{minted}
    \end{minipage}\hfill%
    \begin{minipage}[t]{0.22\textwidth}
        \vspace{0pt} % Add this to ensure top alignment within minipage
        \begin{minted}{julia}
            y .= 0
            for j = _
              let x_j = x[j]
                y_j .= 0
                for i = _
                  let A_ij = A[i, j]
                    y[i] += x_j * A_ij
                    y_j[] += A_ij * x[i]
                  end
                end
                #D is the diagonal
                y[j] += y_j[] + D[j] * x_j
              end
            end
        \end{minted}
    \end{minipage}
    \vspace{-8pt}
    \caption{Finch row-major, column-major and symmetric SpMV Programs}
    \label{spmv_programs}
    \vspace{-12pt}
  \end{wrapfigure}
  \subsection{Sparse Matrix-Vector Multiply (SpMV)}
  Sparse matrix-vector multiplication (SpMV) has a wide range of applications and has been thoroughly studied~\cite{liu_csr5_2015,
  zhou_enabling_2020}. 
  %
  Because SpMV is bandwidth bound, many formats have
  been proposed to reduce the footprint~\cite{langr_evaluation_2016}. 
  %
  The wide range of applications unsurprisingly
  results in a wide range of array structures, making it an effective kernel to
  demonstrate the utility of our programming model. 
  %
  We varied both the data formats and the SpMV algorithm. 
  %
  Our formats
  are shown in Table~\ref{spmv_tensor_formats}, and our programs are listed in
  Figure~\ref{spmv_programs}.
  %
  In varying the program, we consider both row and column-major SpMV programs, as
  well as a symmetric SpMV. 
  %
  Finch enables us to exploit symmetry
  effectively.
  %
  Our program restricts our attention to the canonical triangle (using
  masks), reuses the reads to the canonical triangle of the symmetric matrix
  (using a $\finchdefine$ statement), and writes the results to both relevant
  locations (using multiple outputs).
  %
  All of our programs apply to all of our level formats, enabling exploitation of multiple structural patterns concurrently (e.g. sparsity \textit{and} symmetry).

  Figure~\ref{fig:spmv_grouped} displays speedup relative to TACO, SuiteSparseGraphBLAS, and Julia’s standard library.  We test using sparse matrices from a large selection of datasets spanning several previous papers: the datasets used by Vuduc et al. to test the OSKI interface \cite{vuduc2005oski}, Ahrens et al. to test a variable block row format partitioning strategy \cite{ahrens_optimal_2021}, and Kjolstad et al. to test the TACO library \cite{kjolstad_tensor_2017}. Additionally, we included the SNAP graph collection to test with boolean matrices. We also created some synthetic matrices containing bands of varying sizes.
  % or blocks of varying sizes as well as a permutation matrix to encapsulate a few additional use cases.
  %The dense vector is randomly generated. 
  %
  We tested using the row-major and column-major Finch programs in Figure \ref{spmv_programs} as well as the symmetric program where applicable; the performance displayed for Finch on each dataset is the fastest among the formats and programs we tested. For TACO, Column-major SpMV consistently performs better than row-major SpMV (an average of 1.36x better) so we use column-major SpMV in TACO as our baseline.

  We found that the SpMV performance was superior for the level format that best paralleled the structure of the tensor.
  %
  Namely, matrices with a clear blocked structure like exdata\_1, TSOPF\_RS\_b678\_c1, and heart3 performed notably well with the SparseVBL format with speedups of 2.16, 1.55, and 1.30 relative to TACO, while the baseline format had slowdowns of 0.71, 0.53, and 0.92 relative to TACO.
  %
  Furthermore, the synthetic banded matrices we constructed performed the best with the SparseBand matrix, in particular with the large\_band and the medium\_band matrices having a speedup of 1.98 and 1.64 relative to TACO, while the baseline format had slowdowns of 0.84 and 0.51 relative to TACO.
  %
  There were also significant advantages of using the Pattern format instead of the Element format to represent scalar values in the matrices when these values were boolean, such as matrices in the SNAP collection which represent graph datasets are boolean. 
  %
  For example, the SparseList-Pattern for email-Eu-core resulted in a speedup of 2.51, while the SparseList-Element format resulted in a slowdown of 1.84 over TACO.
  %
  Every symmetric matrix in the SparseList and SparseList-Pattern formats has better performance when we use a Finch SpMV program that takes advantage of this symmetry.
  %
  Symmetric SpMV with the SparseList level format in Finch results in an average of 1.27x speedup over TACO and symmetric SpMV with the SparseList-Pattern format in Finch results in an average speedup of 1.21x over TACO.
  %
  Notably, there is a 1.91x speedup for the HB/saylr4 matrix over TACO.

@@

  % \begin{wrapfigure}{r}{0.41\linewidth}
  %     \begin{minipage}[t]{0.18\textwidth}
  %         \vspace{0pt} % Add this to ensure top alignment within minipage
  %         \begin{minted}{julia}
  %             y .= 0
  %             for j = _, i = _
  %               y[i] += A[i, j] * x[j]
  %             end
  %         \end{minted}
  %         \vspace{24pt} % Add this to ensure top alignment within minipage
  %         \begin{minted}{julia}
  %             y .= 0
  %             for j = _, i = _
  %               y[j] += A[i, j] * x[i]
  %             end
  %         \end{minted}
  %     \end{minipage}\hfill%
  %     \begin{minipage}[t]{0.22\textwidth}
  %         \vspace{0pt} % Add this to ensure top alignment within minipage
  %         \begin{minted}{julia}
  %             y .= 0
  %             for j = _
  %               let x_j = x[j]
  %                 y_j .= 0
  %                 for i = _
  %                   let A_ij = A[i, j]
  %                     y[i] += x_j * A_ij
  %                     y_j[] += A_ij * x[i]
  %                   end
  %                 end
  %                 #D is the diagonal
  %                 y[j] += y_j[] + D[j] * x_j
  %               end
  %             end
  %         \end{minted}
  %     \end{minipage}
  %     \vspace{-8pt}
  %     \caption{Finch row-major, column-major and symmetric SpMV Programs}
  %     \label{spmv_programs}
  %     \vspace{-12pt}
  % \end{wrapfigure}

### 2024-04-05 sections/tensors.tex
Update on Overleaf.

\begin{figure}[htbp]
    \scriptsize
    \raggedright
$declare(lvl, init, dims...)$: Declares the level to hold subtensors
of size $dims$ and an initial value of $init$. Requires the level to be in
read-only mode.

@@

$freeze(lvl)$: Finalizes the updates in the level, and readies the
level for reading. Requires the level to be in update-only mode.

@@

$thaw(lvl)$: Prepares the level to accept updates. Requires the level
to be in read-only mode.

@@

$unfurl(fiber(lvl, pos), [ext], mode, [op], [rhs])$: When $ext$ is given, unfurls the fiber at position
$pos$ in the level $lvl$ over the extent $ext$. When $mode = \finchread$,
returns a looplet nest over the values in the read-only fiber.  When $mode =
\finchupdate$, returns a looplet nest over mutable subfibers in the update-only
fiber. When applied to a scalar or leaf node, $ext$ is omitted and we may also use $op$ and $rhs$ to update the scalar value. Often, skipping over mutable locations allows the level to know which
locations must be stored. Often, a dirty bit may be used in $tns$ to communicate
whether the mutable subfiber has been written to, which allows the parent fiber
to know whether the subfiber must be stored explicitly.

@@

$assemble(lvl, pos_{start}, pos_{stop})$: Allocates subfibers in the
level from positions $pos_{start}$ to $pos_{stop}$. Usually, this function is
only ever called on unassembled positions, but some levels (such as dense levels
or bytemaps) may support reassembly.
\vspace{-8pt}
\caption{The five functions that define a level.}
\vspace{-12pt}
\end{figure}

>>>>

\begin{figure}[htbp]
    \scriptsize
    \raggedright
\fbox{\parbox{\textwidth}{$declare(lvl, init, dims...)$: Declares the level to hold subtensors of size $dims$ and an initial value of $init$. Requires the level to be in
read-only mode.}}

@@

\fbox{\parbox{\textwidth}{$freeze(lvl)$: Finalizes the updates in the level, and readies the
level for reading. Requires the level to be in update-only mode.}}

@@

\fbox{\parbox{\textwidth}{$thaw(lvl)$: Prepares the level to accept updates. Requires the level
to be in read-only mode.}}

@@

\fbox{\parbox{\textwidth}{$unfurl(fiber(lvl, pos), [ext], mode, [op], [rhs])$: When $ext$ is given, unfurls the fiber at position
$pos$ in the level $lvl$ over the extent $ext$. When $mode = \finchread$,
returns a looplet nest over the values in the read-only fiber.  When $mode =
\finchupdate$, returns a looplet nest over mutable subfibers in the update-only
fiber. When applied to a scalar or leaf node, $ext$ is omitted and we may also use $op$ and $rhs$ to update the scalar value. Often, skipping over mutable locations allows the level to know which
locations must be stored. Often, a dirty bit may be used in $tns$ to communicate
whether the mutable subfiber has been written to, which allows the parent fiber
to know whether the subfiber must be stored explicitly.}}

@@

\fbox{\parbox{\textwidth}{$assemble(lvl, pos_{start}, pos_{stop})$: Allocates subfibers in the
level from positions $pos_{start}$ to $pos_{stop}$. Usually, this function is
only ever called on unassembled positions, but some levels (such as dense levels
or bytemaps) may support reassembly.}}
\vspace{-8pt}
\caption{The five functions that define a level.}
\vspace{-12pt}
\end{figure}

\begin{figure}[ht]
\footnotesize
\begin{tabular} {|c|p{0.65\linewidth}|} 
    \hline
    $declare(lvl, init, dims...)$ & Declares the level to hold subtensors of size $dims$ and an initial value of $init$. Requires the level to be in
read-only mode.
\hline

    \end{tabular}

\fbox{\parbox{\textwidth}{$freeze(lvl)$: Finalizes the updates in the level, and readies the
level for reading. Requires the level to be in update-only mode.}}

\fbox{\parbox{\textwidth}{$thaw(lvl)$: Prepares the level to accept updates. Requires the level
to be in read-only mode.}}

\fbox{\parbox{\textwidth}{$unfurl(fiber(lvl, pos), [ext], mode, [op], [rhs])$: When $ext$ is given, unfurls the fiber at position
$pos$ in the level $lvl$ over the extent $ext$. When $mode = \finchread$,
returns a looplet nest over the values in the read-only fiber.  When $mode =
\finchupdate$, returns a looplet nest over mutable subfibers in the update-only
fiber. When applied to a scalar or leaf node, $ext$ is omitted and we may also use $op$ and $rhs$ to update the scalar value. Often, skipping over mutable locations allows the level to know which
locations must be stored. Often, a dirty bit may be used in $tns$ to communicate
whether the mutable subfiber has been written to, which allows the parent fiber
to know whether the subfiber must be stored explicitly.}}

\fbox{\parbox{\textwidth}{$assemble(lvl, pos_{start}, pos_{stop})$: Allocates subfibers in the
level from positions $pos_{start}$ to $pos_{stop}$. Usually, this function is
only ever called on unassembled positions, but some levels (such as dense levels
or bytemaps) may support reassembly.}}
\vspace{-8pt}
\caption{The five functions that define a level.}
\vspace{-12pt}
\end{figure}

% \begin{figure}[htbp]
%     \scriptsize
% \begin{tabular}{|p|p|} 
% \hline
%      Function & Description  \\
%      \hline
%      $declare(lvl, init, dims...)$ & Declares the level to hold subtensors
% of size $dims$ and an initial value of $init$. Requires the level to be in
% read-only mode.\\
% \hline
% $freeze(lvl)$ & Finalizes the updates in the level, and readies the
% level for reading. Requires the level to be in update-only mode.\\
% \hline
% $thaw(lvl)$ & Prepares the level to accept updates. Requires the level
% to be in read-only mode.\\
% \hline
% $unfurl(fiber(lvl, pos), [ext], mode, [op], [rhs])$ & When $ext$ is given, unfurls the fiber at position $pos$ in the level $lvl$ over the extent $ext$. When $mode = \finchread$, returns a looplet nest over the values in the read-only fiber.  When $mode = \finchupdate$, returns a looplet nest over mutable subfibers in the update-only fiber. When applied to a scalar or leaf node, $ext$ is omitted and we may also use $op$ and $rhs$ to update the scalar value. Often, skipping over mutable locations allows the level to know which
% locations must be stored. Often, a dirty bit may be used in $tns$ to communicate
% whether the mutable subfiber has been written to, which allows the parent fiber
% to know whether the subfiber must be stored explicitly.\\
% \hline

% \end{tabular}
% \caption{The five functions that define a level.}
% \vspace{-12pt}
% \end{figure}
